---
title: 'Clinical correlations: anterior vs posterior putamen'
author: "Martin E. Johansson"
date: "2025-06-04"
output: 
  html_document: 
    toc: yes
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Do motor symptoms and their progression depend more on activation in the anterior compared to the posterior putamen?

## Methodology

- Carry out everything up until the 1st-level analysis.
- Extract selection- and motor-related activation from the putamen, parcellated
into anterior, posterior, left, and right. In our case, due to the horizontal
flipping of contrast images, left is most affected whereas right is least
affected.
whereas
- Test whether correlations with motor progression differ for anterior and posterior beta estimates.
- Test for PROGRESSION x ANTERIOR/POSTERIOR x LEFT/RIGHT interactions on 2-year change in beta estimates.
- Also for SEVERITY x ANTERIOR/POSTERIOR x LEFT/RIGHT interactions on beta estimates.

## Load libraries

```{r, echo=F,warning=F,message=F}
library(tidyverse)
library(GGally)
library(mice)
library(cocor)
library(car)
library(lme4)
library(ggeffects)
```

## Load data

```{r, echo=F,warning=F,message=F}
# df_ba_covs <- read_csv('/Volumes/project/3022026.01/pep/ClinVars_10-08-2023/derivatives/merged_manipulated_2024-07-17.csv',
#                        col_select = all_of(c('pseudonym','ParticipantType','TimepointNr','Age',
#                                              'Gender','NpsEducYears','YearsSinceDiag','PrefHand'))) %>%
#   filter(ParticipantType == 'PD_POM',
#          TimepointNr == 0) %>%
#   mutate(Gender=if_else(Gender=='Female',0,1))
# 
# source("/Users/marjoh/Documents/work/code/PPP_imaging/R/functions/retrieve_resphand.R")
# resphand <- retrieve_resphand() %>% rename(ParticipantType=Group) %>% filter(str_detect(Timepoint,'Visit1'))
# df_ba_covs <- df_ba_covs %>%
#   left_join(., resphand, by = c('pseudonym','ParticipantType')) %>%
#   mutate(RespHandIsDominant = if_else(RespondingHand == PrefHand | PrefHand == 'NoPref',1,0)) %>%
#   select(-c(PrefHand,RespondingHand,Timepoint))
# 
# df_clin <- read_csv('/Volumes/project/3022026.01/pep/ClinVars_10-08-2023/derivatives/merged_manipulated_2024-07-17.csv',
#                     col_select = c('pseudonym','ParticipantType','TimepointNr',
#                                    'Up3OnBradySum','z_MoCA__total', 'LEDD')) %>%
#   filter(ParticipantType == 'PD_POM',
#          TimepointNr != 1) %>%
#   pivot_wider(id_cols = c('pseudonym','ParticipantType'),
#               names_from = 'TimepointNr',
#               values_from = c('Up3OnBradySum','z_MoCA__total', 'LEDD'),
#               names_prefix = 'T') %>%
#   mutate(Up3OnBradySum_Delta = Up3OnBradySum_T2 - Up3OnBradySum_T0,
#          z_MoCA__total_Delta = z_MoCA__total_T2 - z_MoCA__total_T0,
#          LEDD_Delta = LEDD_T2 - LEDD_T0)
# # Add covariates and clinical variables
# df_clin <- df_clin %>%
#   left_join(., df_ba_covs[,1:8], by=c('pseudonym','ParticipantType'))
# # Select variables
# df_clin_s <- df_clin %>%
#   select(pseudonym,Age,Gender,NpsEducYears,YearsSinceDiag,RespHandIsDominant,
#          Up3OnBradySum_Delta,Up3OnBradySum_T0,
#          z_MoCA__total_Delta,z_MoCA__total_T0,
#          LEDD_Delta, LEDD_T0)
# # Impute missing values
# isna <- apply(df_clin_s, 2, is.na) %>% colSums()
# missing_perc <- round(isna/nrow(df_clin_s), digits = 3)*100
# print(missing_perc)
# df_clin_s_imp <- df_clin_s %>%
#   mice(m=round(5*missing_perc[names(missing_perc)=='Up3OnBradySum_Delta']),
#        maxit = 10,
#        method='pmm',
#        seed=157,
#        print=FALSE) %>%
#   complete() %>%
#   as_tibble()
# 
# df_putamen <- read_csv('/Volumes/project/3024006.02/Analyses/motor_task/Group/Longitudinal/APvsPP/putamen_activity.csv')
# df <- df_putamen %>%
#   left_join(df_clin_s_imp, by = 'pseudonym') %>%
#   relocate(pseudonym,Age,Gender,NpsEducYears,YearsSinceDiag,
#            RespHandIsDominant,Up3OnBradySum_Delta,Up3OnBradySum_T0,
#            z_MoCA__total_Delta,z_MoCA__total_T0,LEDD_Delta,LEDD_T0,)
# write_csv(df, '/Volumes/project/3024006.02/Analyses/motor_task/Group/Longitudinal/APvsPP/df_combined_clin_putamen.csv')
df <- read_csv('/Volumes/project/3024006.02/Analyses/motor_task/Group/Longitudinal/APvsPP/df_combined_clin_putamen.csv')

df_l <- df %>%
  select(1:12, ends_with('_a'), ends_with('_p'), -contains('_fu')) %>%
  pivot_longer(cols = starts_with('con_'),
               names_to = c('contrast','metric','putamen'),
               names_pattern = '(.*)_(.*)_(.*)') %>%
  rename(beta = value)
df_l2 <- df %>%
  select(1:12, ends_with('_r'), ends_with('_l'), -contains('_fu')) %>%
  pivot_longer(cols = starts_with('con_'),
               names_to = c('contrast','metric','putamen','hemisphere'),
               names_pattern = '(.*)_(.*)_(.*)_(.*)') %>%
  rename(beta = value)
df_l_07 <- df_l2 %>%
  filter(contrast=='con_0007') %>%
  pivot_wider(names_from = metric,
              values_from = beta,
              names_prefix = 'beta_')
df_l_10 <- df_l2 %>%
  filter(contrast=='con_0010') %>%
  pivot_wider(names_from = metric,
              values_from = beta,
              names_prefix = 'beta_')
```

## Analyses

### Change in selection-related activity ~ Motor progression

```{r, echo=F}
# cols <- c('Up3OnBradySum_Delta','z_MoCA__total_Delta','LEDD_Delta',
#           'con_0007_delta_p','con_0007_delta_a',
#           'con_0007_delta_p_r','con_0007_delta_a_r',
#           'con_0007_delta_p_l','con_0007_delta_a_l')
# g <- df %>%
#   select(all_of(cols)) %>%
#   na.omit() %>%
#   ggpairs(lower = list(continuous = wrap('smooth', alpha = 0.5, size = 0.5)),
#           upper = list(continuous = wrap('cor', size = 3))) +
#   theme_bw() +
#   theme(legend.position = 'none')
# print(g)

mm <- lmer(beta_delta ~ Up3OnBradySum_Delta*putamen*hemisphere + Up3OnBradySum_T0 + beta_ba + Age + Gender + NpsEducYears + YearsSinceDiag + (1+putamen+hemisphere|pseudonym),
           control = lmerControl(optimizer = 'bobyqa', optCtrl=list(maxfun=2e5)),
           data = df_l_07)
Anova(mm,type=3) %>% print()
ggemmeans(mm, terms = c('Up3OnBradySum_Delta','putamen','hemisphere')) %>%
  plot() +
  labs(x = 'MDS-UPDRS III bradykinetic-rigid (progression)', y = 'Estimated marginal means') +
  theme_bw()
```

### Selection-related activity ~ Motor severity

```{r, echo=F}
# cols <- c('Up3OnBradySum_T0','z_MoCA__total_T0','LEDD_T0',
#           'con_0007_ba_p','con_0007_ba_a',
#           'con_0007_ba_p_r','con_0007_ba_a_r',
#           'con_0007_ba_p_l','con_0007_ba_a_l')
# g <- df %>%
#   select(all_of(cols)) %>%
#   na.omit() %>%
#   ggpairs(lower = list(continuous = wrap('smooth', alpha = 0.5, size = 0.5)),
#           upper = list(continuous = wrap('cor', size = 3))) +
#   theme_bw() +
#   theme(legend.position = 'none')
# print(g)

mm <- lmer(beta_ba ~ Up3OnBradySum_T0*putamen*hemisphere + Age + Gender + NpsEducYears + YearsSinceDiag + (1+putamen+hemisphere|pseudonym),
           control = lmerControl(optimizer = 'bobyqa', optCtrl=list(maxfun=2e5)),
           data = df_l_07)
Anova(mm,type=3) %>% print()
ggemmeans(mm, terms = c('Up3OnBradySum_T0','putamen','hemisphere')) %>%
  plot() +
  labs(x = 'MDS-UPDRS III bradykinetic-rigid (baseline)', y = 'Estimated marginal means') +
  theme_bw()
```

### Change in motor-related activity ~ Motor progression

```{r, echo=F}
# cols <- c('Up3OnBradySum_Delta','z_MoCA__total_Delta','LEDD_Delta',
#           'con_0010_delta_p','con_0010_delta_a',
#           'con_0010_delta_p_r','con_0010_delta_a_r',
#           'con_0010_delta_p_l','con_0010_delta_a_l')
# g <- df %>%
#   select(all_of(cols)) %>%
#   na.omit() %>%
#   ggpairs(lower = list(continuous = wrap('smooth', alpha = 0.5, size = 0.5)),
#           upper = list(continuous = wrap('cor', size = 3))) +
#   theme_bw() +
#   theme(legend.position = 'none')
# print(g)

mm <- lmer(beta_delta ~ Up3OnBradySum_Delta*putamen*hemisphere + Up3OnBradySum_T0 + beta_ba + Age + Gender + NpsEducYears + YearsSinceDiag + (1+putamen|pseudonym),
           control = lmerControl(optimizer = 'bobyqa', optCtrl=list(maxfun=2e5)),
           data = df_l_10)
Anova(mm,type=3) %>% print()
ggemmeans(mm, terms = c('Up3OnBradySum_Delta','putamen','hemisphere')) %>%
  plot() +
  labs(x = 'MDS-UPDRS III bradykinetic-rigid', y = 'Estimated marginal means') +
  theme_bw()
```

### Motor-related activity ~ Motor severity

```{r, echo=F}
# cols <- c('Up3OnBradySum_T0','z_MoCA__total_T0','LEDD_T0',
#           'con_0010_ba_p','con_0010_ba_a',
#           'con_0010_ba_p_r','con_0010_ba_a_r',
#           'con_0010_ba_p_l','con_0010_ba_a_l')
# g <- df %>%
#   select(all_of(cols)) %>%
#   na.omit() %>%
#   ggpairs(lower = list(continuous = wrap('smooth', alpha = 0.5, size = 0.5)),
#           upper = list(continuous = wrap('cor', size = 3))) +
#   theme_bw() +
#   theme(legend.position = 'none')
# print(g)

mm <- lmer(beta_ba ~ Up3OnBradySum_T0*putamen*hemisphere + Age + Gender + NpsEducYears + YearsSinceDiag + (1+putamen+hemisphere|pseudonym),
           control = lmerControl(optimizer = 'bobyqa', optCtrl=list(maxfun=2e5)),
           data = df_l_10)
Anova(mm,type=3) %>% print()
ggemmeans(mm, terms = c('Up3OnBradySum_T0','putamen','hemisphere')) %>%
  plot() +
  labs(x = 'MDS-UPDRS III bradykinetic-rigid (baseline)', y = 'Estimated marginal means') +
  theme_bw()
```
